% !TEX root = thesis-main.tex

\section{Origins}
\label{section:introduction:origins}

The research presented in Chapters~\ref{chapter:research-01}-\ref{chapter:research-05} of this thesis is based on a number of peer-reviewed publications. 
Below, we indicate the origins of each chapter.

\paragraph{Chapter~\ref{chapter:research-01}} is based on Marzieh Fadaee and Arianna Bisazza and Christof Monz,
``Learning Topic-Sensitive Word Representations'',
\textit{In Proceedings of the 55th Annual Meeting of the Association for Computational Linguistics (ACL)},
%pages 441--447,
%Vancouver, Canada,
%July 2017.
\citep{fadaee-etal-2017-learning}.
Fadaee designed and carried out the experiments. All authors contributed to the discussion and text.

\paragraph{Chapter~\ref{chapter:research-02}} is based on Marzieh Fadaee and Arianna Bisazza and Christof Monz,
``Data Augmentation for Low-Resource Neural Machine Translation'',
\textit{In Proceedings of the 55th Annual Meeting of the Association for Computational Linguistics (ACL)},
%pages 567--573,
%Vancouver, Canada,
%July 2017.
 \citep{fadaee-bisazza-monz:2017:Short2}.
Fadaee designed the methods, performed the experiments and wrote most of the text.
Bisazza and Monz contributed to the discussion and editing.

\paragraph{Chapter~\ref{chapter:research-03}} is based on Marzieh Fadaee and Christof Monz,
``Back-Translation Sampling by Targeting Difficult Words in Neural Machine Translation'',
\textit{In Proceedings of the 2018 Conference on Empirical Methods in Natural Language Processing (EMNLP)},
%pages 436--446,
%Brussels, Belgium,
%October 2018.
\citep{fadaee-monz-2018-back}.
Fadaee designed the methods, performed the experiments and wrote most of the text.
Monz contributed to the discussion and editing.

\paragraph{Chapter~\ref{chapter:research-04}} is based on Marzieh Fadaee and Arianna Bisazza and Christof Monz,
``Examining the Tip of the Iceberg: A Data Set for Idiom Translation'',
\textit{In Proceedings of the Ninth International Conference on Language Resources and Evaluation (LREC)},
%pages 925--929,
%Miyazaki, Japan,
%May 2018.
\citep{L18-1148}.
Fadaee designed the methods, performed the experiments, and wrote the text.
Bisazza and Monz contributed to the discussion and editing.

\paragraph{Chapter~\ref{chapter:research-05}} is based on Marzieh Fadaee and Christof Monz,
``The Unreasonable Volatility of Neural Machine Translation'', 
\textit{In Proceedings of the 4th Workshop on Neural Generation and Translation (WNGT)},
%Seattle, Washington, USA.
%July 2020.
\citep{fadaee_new}.
Fadaee designed the methods, performed the experiments, and wrote the text.
Monz contributed to the discussion and editing.



