\blankpage

{\pagestyle{empty}

{
\begin{center}

\noindent
\textbf{Acknowledgements} \\ \vskip .5cm
\end{center}
}

\begin{quote}
{``In the good mystery there is nothing wasted, no sentence, no word that is not significant. And even if it is not significant, it has the potential to be so - which amounts to the same thing.''  \footnote{Paul Auster, The New York Trilogy}}
\end{quote}
\noindent 
\textit{No word that is not significant}; this is the main discussion of this book on the importance of context as well as an accurate description of how I (and this book) came to be. 
%rewrite ^
When I started my PhD the world was a different place. Then again, that can be said about every minute that goes by. What remains a constant is the impact of people around us, in the beginning of things and in the end.
It has been an agonizingly fun ride! And I am very fortunate to have many amazing people around me that made it possible. 

%christof
First, Christof.
I came to Amsterdam with a promise and it turned out to be quite an experience. 
I owe my deepest gratitude to Christof for trusting me and taking me on this journey. 
Over the years he taught me how to research new areas, how to break down complicated questions to get to the heart of the problem, and how to never be happy with an answer and always push for more.
I specially would like to thank him for his attentiveness when it counted.
I enjoyed the walks, the talks, and making him exasperated with my wordplays. 

%arianna
I would like to thank Arianna for her guidance and for being a friend. 
Her critical thinking and questioning the things that I took for granted taught me how to explore novel corners of research and ask new questions. 
I will always be grateful to have her by my side, whether it was during discussions with Christof or the disappointment that was the Star Wars sequels.

%maarten
Next, I would like to express my gratitude to Maarten for building a great research group at ILPS. 
From the first day of joining ILPS, I felt part of a team. I enjoyed the fun impromptu discussions, the group activities, and everything in between.
Now that I've been away for some time, I see that what we had at ILPS was something truly special and Maarten is a big reason for that. 
Thanks for the guidance and for encouraging me to push my boundaries. 

%committee
My sincere gratitude to Alexandra, Antal, Paul, Evangelos, and Ilya who were generous with their time to read this thesis and accepted to be part of my defense committee. 

%mt
I would like to thank the MT members: Marlies, Ke, Katya, and Praveen, my brothers and sisters in arms. 
We shared laughs and tears, successes and failures, valuable and worthless discussions. And we all came out of it alive! 
What I got out of this PhD was more than I expected and I am glad to have met these people along the way. 
Thanks for making my PhD experience a fun and enjoyable one. 

%ilps
I am grateful for many current and former members at ILPS for our times together. From showing the ropes and sharing their wisdom, to asking intriguing questions, to fun sushi parties. 
Thanks Alexey, Amir, Anna, Artem, Boris, Chang, Chuan, Daan, Damien, Dan, David, David, Dilek, Eva, Evgeny, Fei, Harrie, Hendra, Hendrike, Hinda, Ivan, Jiahuan, Jie, Julia, Julien, Kaspar, Maarten, Maurits, Mostafa, Pengjie, Richard, Ridho, Rolf, Shangsong, Shaojie, Spyretta, Svitlana, Tobias, Trond, Vera, Xiaohui, Xiaojuan, Xinyi, Yangjun, Yifan, Zhaochun, and Ziming.

I particularly like to thank Anne, Ana, Bob, Christophe, Evangelos, Ilya, Isaac, Manos, and Tom, who are responsible for some of my good memories.
Special thanks to Petra, who knows the answer to every question and is a delight to be around. 
%
%cover
I would also like to thank Mostafa for designing the cover and putting up with all of my typography suggestions.
%et.
Cristina, thanks for being a ray of sunshine full of support and insight. We will always have Toronto.

%paranymphs
Next, I want to thank my wonderful paranymphs: Maartje, my office-mate and friend who makes an excellent arrabiata pasta, and Nikos, my partner in crime, who makes every moment the opposite of boring. 
Thanks for agreeing to be by my side on my defense day.

%ebay
I spent a summer in Aachen doing an internship at eBay which was an invaluable practice. Thanks, Shahram and Jose for the guidance and the interesting discussions, and Daniel, Leonard, Michael, Nicola, and Sivan for making it a nice experience. 
Parnia thanks for the talks, the hikes, and the Ash Reshtehs. 

%nl friends
Moving to Amsterdam meant a fresh start and a new beginning. I was lucky to meet many amazing people that help me weather the gray days and savor the sunny ones. I am particularly thankful to Ali, Ali, Ameneh, Amin, Amirhosein, Arash, Aylar, Azad, Azadeh, Behrouz, Danial, Fahimeh, Fatemeh, Hester, Hoda, Hoda, Hojat, Hoora, Irene, Iris, Jasmijn, Jeyran, Mahdi, Mahdi, Mahdieh, Maria, Maryam, Marzieh, Marzieh, Masoud, Masoumeh, Mohammad, Mohammad Amin, Mohammad Hosein, Mohsen, Narges, Naser, Nasim, Nasrin, Parisa, Parisa, Parisa, Saba, Samira, Samira, Sara, Shayan, Shima, Vahid, Vivianne, Wilker, Zahra, and Zoheir.   
A particular thanks to a special group of people with whom I had many imaginative conversations: Ava, Behsa, Borna, NikAyeen, Pouya, Radin, Saba, Samin, Sarina, Sepehr, Sina, and Zeinab. They made me an optimist about the future. 


%algon
My Algonquin Round Table, whom I am grateful to have in my life:
Abbas, Ali, Arian, Atieh, Elaheh, Faezeh, Farzad, Ghazaleh, Louise, Mandana, Maryam, Mehdi, Safoora, Sara, Sara, Siavash. 
I met each of these people at different points in my life and something about each of them got stuck with me. Geographical determinism might be working against us now, but I cherish every time we got a chance to meet somewhere in this world. 

%bros
My brothers Mohammad and Mohsen, thanks for doing what brothers do best: supporting me in times of need and humbling me the rest of the time. Fatemeh and Mahsa, I am thankful for the kind words and the support.

%parents
I am eternally grateful to my amazing parents. 
Baba, you showed me every day that there is always something new worth learning. 
Your dedication to diving blindly into new areas, languages, and technologies is always inspiring.
You taught me to never stop fighting for what I want. 
Maman, you are the strongest woman I know and I appreciate everything you have done for me.
Every step I took has been a little bit easier because of the hard decisions you had to make. 
I followed my dreams because of you.


%hamid
Lastly, Hamid. 
You are the kindest, smartest, most driven person I know. 
And yes I am still not a fan of using superlatives. 
It is truly earned in your case though.
This would not have been possible without you. 
You were the best sounding board for everything that did (and did not) end up in this book.
I am lucky to have your love, your support, and your wisdom by my side in this journey. 
\\
\\
\grade{Marzieh}
\grade{September 2020}

\vfill

\begin{quote}
{``Everything becomes essence; the center of the book shifts with each event that propels it forward. The center, then, is everywhere, and no circumference can be drawn until the book has come to its end.'' \footnote{Paul Auster, The New York Trilogy}}
\end{quote}

\clearpage
}

\blankpage
